---
title: "Cherry_Blossum"
author: "RLT"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# For the Story of Cherry Blossum see

https://edition.cnn.com/2021/04/05/asia/japan-cherry-blossoms-climate-change-intl-hnk-scn/index.html


```{r}
library(tidyverse)
```

See the following link for an analysis of the time of blossom of the Cherry Blossom trees in Tokyo, Japan from year 801 to this year.

The objective of this excecise to to put in practice the functions different files and try and replicate some of the figures that have been published in newspapers.   

See the following websites
1. 
https://edition.cnn.com/2021/04/05/asia/japan-cherry-blossoms-climate-change-intl-hnk-scn/index.html

2. 

https://www.yahoo.com/gma/numbers-japan-cherry-blossoms-hit-214535221.html

3. 

https://www.meteored.com.ar/noticias/actualidad/sorpresa-en-japon-por-la-temprana-floracion-de-los-cerezos.html



You will find 4 original files from the authors with the references, ect.

1. 759TempW.xls
2. KyotoFullFlowerW.xls
3. TempReconsWFinal.xls
4. CO2_r0.csv


# Step 1.

a. Remove all unnesecary information for each Excel file
b. Convert files as .csv

# Step 2
Load files in a RMarkdown document (make certain they are readible)

# Step 3 
## (5 points)

Merge all three file in one file, "show script and the last 6 lines" (make it clean, remove unnessary information, not the rows but the unnessary columns)
 a. Show the names of tghe columns in the new dataframe

# Step 4 
## (5 points)
  Replace all -999.9 and -99.9 with NA within R.  (Do not do this in Excel, show the script):
 a. show the head and tail of the new data frame.
  
  Use the following website to find an appropriate script
  https://cran.r-project.org/web/packages/naniar/vignettes/replace-with-na.html
  

# Step 5 
## (3 and 2 points)
 a. Calculate the running average of the flowering time for 10 years prior to time t. 
 b. Show the last 10 lines of the data frame


# Step 6 
## (3 points)

a. Create a point diagram of the date (x axis) and time of "Full Flowering date DOY" 

# Step 7 
## (3 and 2 points)

 a. Add a linear regression to the above figure
 b. Does this show a pattern of earlier flowering time?
  
# Step 8 
## (2, 3 and 2 points)

a. Now redo the analysis using an analysis called LOESS regression, this is the default geom_smooth model, geom_smooth(). 
b. Explain what LOESS regression model does?  
c. Now does this show a pattern of change and flowering period?

# Step 9
## (3 and 2 points)

Using the mean "running average of the day of blossom" you calculated above. Use 10 year running average.
 a. Now plot the the rolling average of Blossom Day (10 year average) on your graph.
 b. Porque la linea no se extiende antes del año 1300? comparando al step 7.

# Step 10 
## Bonus (5 points)

a. Join the file with the CO2 data and do an analysis to evaluate the relationship between Average temperature in Tokyo and average CO2 (Rolling average of 10 years). 

 b. Explain how your hypothesis and how this may support or not evidence for climate change





# Imported Files
```{r}
library(readr)
X759TempW_mod <- read_csv("Cherry_Blossom/759TempW_mod.csv")
F1=X759TempW_mod
head(F1)

KyotoFullFlowerW_mod <- read_csv("Cherry_Blossom/KyotoFullFlowerW_mod.csv")
F2=KyotoFullFlowerW_mod
head(F2)

TempReconstWFinal_mod <- read_csv("Cherry_Blossom/TempReconstWFinal_mod.csv")
F3=TempReconstWFinal_mod
head(F3)

```

Join the three files and show the head and tail of the new data frame


```{r}
Cherry_B=F1 %>% 
full_join(F2) %>% 
  full_join(F3)


```

```{r}
Cherry_B
```

# Remove the -999.9 and -99.9.
 and show the head and tail of the new data frame

```{r}
library(naniar)
CB=Cherry_B %>% 
replace_with_na(replace = list(TEMP = c(-999.9),
                             Obs_Temp = c(-999.9),
                             Temp_3_31= c(-99.99),
                             Temp_3_31_1 = c(-99.99),
                             TemP_3_22 = c(-99.99),
                             Conf_CI_low = c(-99.99),
                             Conf_CI_High = c(-99.99)))
head(CB)
```


# Create a point diagram of the date (x axis) and time of "Full Flowering date DOY" 


```{r}
ggplot(CB, aes(AD, Full_flow_date_DOY))+
  geom_point()
```

# Add a linear regression to the above figure
  Does this show a pattern of earlier flowering time?

```{r}
ggplot(CB, aes(AD, Full_flow_date_DOY))+
  geom_point()+
  geom_smooth(method="lm")
```



Now redo the analysis using an analysis of LOESS, the default geom_smooth model. Explain hat LOESS egression model does?  Now does this show a pattern of change and flowering period?


```{r}
ggplot(CB, aes(AD, Full_flow_date_DOY))+
  geom_point()+
  geom_smooth()
```



#  Calculate the mean "running average of the day of blossom". Use 10 year running average.


```{r}
names(CB)
library(zoo)
CB=CB %>% 
  mutate(RAB = rollmean(Full_flow_date_DOY, k = 10, fill = NA))
```

# Now plot the the rolling Blossom Day (10 year average) on your graph

 Por que la linea no se extiende antes del año 1300

```{r}
ggplot(CB, aes(AD, Full_flow_date_DOY))+
  geom_point()+
  geom_smooth(aes(AD, RAB))
```



```{r}
library(readr)
CO2_Yr0 <- read_csv("Cherry_Blossom/CO2_Yr0.csv")


CB=CB %>% 
  right_join(CO2_Yr0 , c("AD" = "year"))
  
```


Do a figure showing the relationship between mean global Co2 concentration and time of flowering. What does the patern show?.  

```{r}
ggplot(CB, aes(data_mean_global_CO2, Full_flow_date_DOY))+
  geom_point()+
  geom_smooth()
```


Now do a figure showing the mean average temperature (10 year running average) and C02 concentration. To calculate the mean running average of the temparature use the following variable, Temp_3_31_1. (which is the mean march temperature). 

```{r}
CB
```


```{r}
names(CB)
library(zoo)
CB=CB %>% 
  mutate(RAT = rollmean(Temp_3_31_1, k = 10, fill = NA))


ggplot(CB, aes(data_mean_global_CO2, RAT))+
  geom_point()+
  geom_smooth()


```

