---
title: "Data_COVID_PR"
subtitle: "BIOL4026"
author: "Raymond L. Tremblay"

output:
  html_document:
    css: tutorial.css
    fig_caption: yes
    highlight: pygments
#    theme: simplex
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

```{=html}
<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #1125d9;
}
</style>
```


```{r echo=FALSE}

Sys.Date()

rm(list = ls()) # Remove information from the global environment
```

***
## Enlace de Monitoreo de COVID-19 en Puerto Rico

Para ver analysis de COVID -19 en PR vea este enlace

https://rconnect.dfci.harvard.edu/covidpr/





*** 
COVID_PR   This is the script to get access to the most recent file of the COVID cases in PR

```{r }
library(data.table)
url_COVID_PR=fread("https://raw.githubusercontent.com/rafalab/pr-covid/master/dashboard/data/DatosMortalidad.csv")

```


```{r}
## This is too save the file to your computer
#head(url_COVID_PR)
library(utils)
#write.csv(url_COVID_PR, "Datos/url_COVID_PR.csv")
```

# Upload file from your computer (you may not need this)

```{r}
library(readr)
url_COVID_PR <- read_csv("Datos/url_COVID_PR.csv")
```

### Look at your data

Understand what each column means

```{r}


names(url_COVID_PR)
head(url_COVID_PR)
tail(url_COVID_PR, n=10)

```

***

## Lubridate package

## GET YOU CHEAT SHEET HERE

https://rstudio.com/resources/cheatsheets/



Separate the date in Years/Month/Day, each in their own column

Use the library(lubridate)

```{r}
library(tidyverse)
library(lubridate)
```

***


```{r}


url_COVID_PR$mdy=mdy(url_COVID_PR$Fecha)
head(url_COVID_PR)
```

***
## Extract Date

Extract day, month and year in different column.
Let us start with extracting the month for year data point

```{r}
url_COVID_PR$month=month(url_COVID_PR$mdy)

tail(url_COVID_PR)



```

## Now extract "year" and the "day" in different columns


```{r}

url_COVID_PR$day=day(url_COVID_PR$mdy)
url_COVID_PR$year=year(url_COVID_PR$mdy)

head(url_COVID_PR)
```




## Now extract the day of the week with the function "wday".  Which number is Monday?

```{r}
url_COVID_PR$weekday=wday(url_COVID_PR$mdy)

tail(url_COVID_PR)
```

***

## Tarea.

1. Find the day of where there was the most number of NEW COVID cases in a specific day. ( 5 points)

2. Find the day of where there was the most number of New Deaths in a specific day. ( 5 points)

3. Find the day where there was the most number of beds of ICU (Intensive Care Units) that were occupied "CamasICU". ( 5 points)

4. Calculate the total mean number of deaths for each month. ( 8 points)

5. Do a point graph of the mean number of deaths per month (y axis) and the months surveyed (x-axis) ( 5 points)

6. Which day of the week is there the largest mean number of new COVID cases.  Do a figure a 7 days of the weeks and the mean number of new cases (7 points) 


```{r}
url_COVID_PR %>% 
  dplyr::select(weekday, IncCasosSaludNuevo) %>% 
  group_by(weekday) %>% 
  summarise(mean_New_cases=mean(IncCasosSaludNuevo, na.rm=TRUE))
```


***

Submission of work

1. You will upload your file to Edmodo as a .html file.
2. Make certain you document is clean (non long tables, easy to understand). 
3. Answer every question. 

Date of Submission of your complete work 21st of February at 11:59pm.




